### 2008-08-16 doc/a-rex/arex_tech_doc.tex
More changes to A-REX doc.

\begin{center}
\textsc{\large The ARC Computational Job Management Component - A-Rex}
\par\end{center}{\large \par}

@@

A-Rex is one of ARC middleware componets implementing functions of
so called \emph{Computing Element} (CE).

The aim of the A-Rex is to take care of job pre- and post-processing.
It provides functionality to stage-in files containing input data
and program modules from wide range of sources and transfer or store
output results.

The A-Rex implements Web Service (WS) interface which provides a way
to submit and control computational tasks (jobs) to be executed by
the A-Rex and undelying Local Resource Management System.

@@

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files in the SD as well. The A-Rex does not guarantee the
availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environement).
The SD is also the only place which is controlled by the A-Rex. It
is accessible by the user from outside through HTTP(S) protocol. Any
file created outside the SD is not controlled by the A-Rex. Any exchange
of data between client and A-Rex (including also program modules)
is done through HTTP(S) protocol. A URL for accessing input/output
files is  obtained through the WS Local Information Description Interface
(LIDI) of A-Rex.

Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
\cite{jsdl} XML document - which identifies the job in the A-Rex
and the Information Interface.

Jobs are initiated and controlled through WS interface. Complete job
descriptions (JD)are passed to the A-Rex through WS in JSDL\cite{jsdl}
coded description. Input data files and job executables are transferred
separately through the same interface, as described in the next section.

@@

One of the most important tasks of the A-Rex is to take care of processing
input and output data (files) of the job. Input files are gathered
in the SD or associated cache area. There are two ways to put a file
into the SD:

\begin{itemize}
\item Download is initiated by the A-Rex -- This is the case for files defined
in the JD (with name and source). The A-Rex alone is responsible to
make sure that all required files will be available in the SD.\\
The supported protocols for sources at the moment are (in case of
full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
some less standard sources are supported. Those are described below.
The A-Rex fully relies on HED framework \cite{hed} for data transfering
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-Rex. Examples of such
files are the main executable of the job, the job's input files, etc.
These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-Rex. Access to AFS, NFS, FTP, HTTP and
any other remote data transport during execution of a job is not guaranteed
(at least not by A-Rex).

@@

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service} (RLS) -- The A-Rex takes care
of these files. They have to be specified in the JD. If the job fails
during any stage of processing, no attept is made to transfer those
files to their final destination, unless the option \emph{preserve=yes}
is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}
All files not specified in the JD are deleted after job execution
finished. If job execution fails for any reason (if exit code of main
executable is not 0) all files from furst group are transfered to
second one.

@@

From the point of view of the A-Rex a job passes through various states.
Picture \ref{job states diagram} presents a diagram of the possible
states of a job.%
\begin{figure}[tbh]
\begin{centering}
\includegraphics{pic1}
\par\end{centering}

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the dedicated Local
Information Description Interface of A-Rex using the UI or any other
suitable tool or through query method of WS interface.

@@

Below is the description of all actions taken by the A-Rex at every
state:

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to a
CE but is not processed yet. The A-Rex will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The A-Rex is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If all files
are successfully gathered the job moves to the next state. If \textbf{any}
file can't be downloaded or it takes the UI too long to upload a file,
the job moves to \textbf{Finishing} state. It is possible to put a
limit on the number of simultaneous \textbf{Preparing} jobs. If this
limit is exceeded, jobs ready to enter the \textbf{Preparing} state
 will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
mark. Exceptions are jobs which have no files to be downloaded. These
are processed out of limits.
\item \textbf{Submiting} - The job is being passed for execution to the
\emph{Local Resource Management System} (LRMS). The correspoding backends
for many LRMSs are provided with the default installation. If the
local job submission is successful the job moves to the \textbf{Executing}
state. Otherwise it moves to \textbf{Finishing}. It is possible to
limit the agregate number of jobs in \textbf{Submiting} and \textbf{Executing}
states.
\item \textbf{Executing} - The job is queued or being executed in the LRMS.
The A-Rex takes no actions except waiting until the job finishes.
\item \textbf{Killing} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at an Indexing Service. The user can download data files
from the SD by using the UI or other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the A-Rex.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the Informatrion Interface of the A-Rex. If a
job was moved to \textbf{Finished} because of failure, it may be restarted
on request of a client. When restarted, a job is moved to the state
previous to the one in which it failed and is assigned mark PENDING.
This is needed in order to not break the configuration limits. Exception
is a job failed in \textbf{Executing} state and lacking input files
specified in JD. Such a job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by
the user}. No files will be moved to their destination SE.

@@

In full installation the A-Rex and it's components support the following
data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
http, https, lfc} and \emph{rls.} For more information please see
{}``The Hosting Environment of the Advanced Resource Connector middleware''
document \cite{hed}.

@@

\subsection{Internal Files of the A-Rex}

For each local UNIX user listed in the A-Rex configuration - including
a generic one which covers all local user identities - a \textit{control
directory} exists. In this directory the A-Rex stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. In most common configuration case A-Rex
serves all users defined by Operating System and stores their control
file in same directory. To make it easier to recover in case of failure,
the A-Rex stores most information in files rather than in memory.
All files belonging to the same job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.description} - contains the description of the job
(JD).
\item \textit{job.ID.local} - information about the job used by the A-Rex.
It consists of lines of format \textit{{}``name = value''}. Not
all of them are always available. The following names are defined:

@@

\item \textit{job.ID.proxy} -delegated X509 credentials.
\item \textit{job.ID.proxy.tmp} - temporary X509 credentials with different
UNIX ownership used by processes run with effective \emph{user id}
different form job owner's \emph{id}.
\end{itemize}
There are other files with names like job.ID.{*} which are created
and used by different parts of the A-Rex. Their presence in the \textit{control
directory} can not be guaranteed and can change depending on changes
in the A-Rex code.

@@

The job submission and control interface is based on document produced
by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
The exchange of SOAP messages is performed over the HTTP(S) protocol.
The BES interface is represented by two port-types - BES-Management
and BES-Factory. The former is made to control the A-REX service itself
and thus defines operations to start and stop the functionality of
the BES service. The A-Rex does not implement remote control of service
functionality. Hence the BES-Management port-type is not functional.
The BES-Factory port-type provides operations to submit new jobs (to
create an activity in terms of BES) and to monitor its state. It also
has an ability to provide information about the service. A-Rex fully
implements the functionality of this port-type.

For job descriptions A-Rex accepts the Job Submission Description
Language (JSDL)\cite{jsdl} documents as defined by the OGF Job Submission
Description Language Working Group. Supported elements and extensions
are described below.

@@

A-Rex introduces a new operation in addition to those provided by
BES. It does that by defining its own port-type with the single operation
\emph{ChangeActivityStatus} (see Appendix D). This operation provides
a way to request simple transfers between states of jobs and corresponding
actions.

@@

Since BES allows implementators to extend their initial activity states
with additional sub-states, A-Rex defines a set of sub-states of activity
processing in addition to those defined by the BES, as listed in Table
\ref{tab:Job-states-definitions}. Their meaning is described in section
\ref{Section: Job Flow}.

@@

The A-Rex also supportd the Delegation Interface (see Appendix E).
This is a common purpose interface to be used by ARC services which
accepts delegated credentials from clients. The Delegation Interface
implements two operations: initiatialization of credentials delegation
(DelegateCredentialsInit) and update/renewal of credentials (UpdateCredentials).

@@

\begin{itemize}
\item \emph{UnsupportedFault}: Indicates that service does not support this
operation despite supporting the port-type.
\item \emph{ProcessingFault}: Internal problems during generation of the
token.
\end{itemize}
\end{itemize}
\end{itemize}
Additionally, A-Rex Web Service Interface allows delegation to be
performed as part of the \emph{CreateActivity} operation of the BES-Factory
port-type. For this it accepts the element \emph{DelegatedCredentials}
inside the \emph{CreateActivity} element. The \emph{Id} element of
\emph{DelegatedCredentials} must contain an identifier obtained in
response to the previous \emph{DelegateCredentialsInit} operation.
For more information about delegations and delegation interface refer
to \cite{wsrf-rp}.

@@

The A-REX implements the Local Information Description Interface (LIDI)
interface common for all ARC services. This interface is based on
OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
Information about resources and maintained activities/jobs are represented
in a \emph{WS-Resource Properties} informational XML document. The
document type is defined in the A-Rex WSDL as a \emph{ResourceInformationDocumentType}.
It contains the following elements/resources:

@@

A-REX supports the following elements from the JSDL version 1.0 specification\cite{jsdl}
including POSIX Applications extension:

\begin{lyxlist}{00.00.0000}
\item [{\emph{JobName}}] \emph{-} name of the job as assigned by the user.
\item [{\emph{Executable}\ (POSIX)}] - name of the executable file.
\item [{\emph{Argument}\ (POSIX)}] - arguments the executable will be
launched with.
\item [{\emph{DataStaging}}]~

\begin{lyxlist}{00.00.0000}
\item [{Filename}] - name of the data file on the executing node.
\item [{\emph{Source}}] - source where the file will be taken from before
execution.
\item [{\emph{Target}}] \emph{-} destination the file will be delivered
to after execution.
\end{lyxlist}
\item [{\emph{Input}\ (POSIX)}] - file to be used as standard input for
the executable.
\item [{\emph{Output}\ (POSIX)}] - file to be used as standard output
for the executable.
\item [{\emph{Error}\ (POSIX)}] - file to be used as standard error for
the executable.
\item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
for execution.
\item [{\emph{TotalPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{IndividualPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{CPUTimeLimit}\ (POSIX)}] - maximal amount of CPU time needed
for execution.
\item [{\emph{TotalCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{IndividualCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{WallTimeLimit}\ (POSIX)}] - amount of clock time needed
for execution.
\item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
\item [{\emph{IndividualCPUCount}}] - same as TotalCPUCount.
\end{lyxlist}

@@

The GM can cache input files. Caching is enabled if corresponding
command is present in configuration file. The GM does not cache files
marked as executable in job. Caching can also be explicitly turned
off by user for each file by using \textit{cache=no} option in URL
(for URL options read {}``Protocols, Uniform Resource Locators (URL)
and extensions supported in ARC'' \cite{urls}). The disc space occupied
by cache is controlled by removing unused files. For more information
look in section \ref{SubSection: ConfigFile}.

@@

Before downloading a file the GM tries to determine it's size and
then tries to preallocate space in the cache directory by writing
a file of the same size. If this fails (possibly because the file
system has no more space), it tries to the remove oldest cached files
which are not being used by any job. This means that \textbf{the hard
limit of cache size is the space available in the file-system}. In
case cache gets full and it is impossible to free enough space, the
download fails and then is retried without using cache.

Before giving access to an already cached file, the GM contacts the
original file source to check if the user has suficient access rights
to the file. Not all protocols make this check possible. Also, file
creation or validity times are checked to make sure the cached file
is fresh. If it is impossible to obtain creation and invalidation
times, the file is invalidated 24 hours after it was downloaded.

The GM checks the cache periodically. If the space used by cache exceeds
the high water-mark given in the configuration file (\emph{softsize}),
it tries to remove the oldest unused files until cache size drops
below the low water-mark. This sets the soft limit of cache size.

There are two kinds of caches supported. Files in \textit{private}
cache are owned by the Unix user to which a grid user is mapped. Those
files are readable only by that particular Unix user. Another kind
of cache is \textit{shared}. Files in the shared cache are owned by
the Unix user who started GM and are readable by everyone.

@@

The GM consists of several separate executable modules. These are:

\begin{itemize}
\item \textit{grid-manager} - The main module. It is responsible for processing
jobs, moving them through states and running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at an Indexding Service
(like RLS) as needed. It processes and updates the \textit{job.ID.output}
file.
\item \emph{cache-register} - Utility to register cached data into an Indexing
Service. It reads and modifies cache informational files \emph{old}
and \emph{new} (as described in section \ref{sec:cache}). Configuration
is read directly from the GM's configuration file (see section \ref{SubSection: ConfigFile}).
It is run by the GM every 5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
the frontend. It puts collected information into the \emph{job.ID.diag}
file.
\item \emph{gm-kick} - Sends a signal to the GM though a FIFO file to wake
it up. It's used to increase responsiveness of GM.
\end{itemize}
The following modules are always run under the Unix account to which
a grid user is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Curently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
pseudo-LRMS is supported for testing purposes. This module is responsible
for job submission to the LRMS.
\item \textit{cancel-{*}-job} - This script is for canceling jobs which
have been already submitted to the LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the GM about completion of jobs. It's implementation for PBS uses
server logs to extract information about jobs. If logs are not available
it uses the less reliable \emph{qstat} command for that. Other backends
use different techniques.
\end{itemize}
In addition, there is also an administration utility:

\begin{itemize}
\item \textit{gm-jobs} - prints a list of jobs available on the cluster
and the number of jobs in each state.\\
\hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name]\\
\hspace*{0.5cm}-l -- print more information about each job,\\
\hspace*{0.5cm}-u -- pretend utility is run by user with id \emph{uid},\\
\hspace*{0.5cm}-U -- pretend utility is run by user with name \emph{name}.
\end{itemize}
GM comes with plugins useable for various authorization purposes (see
for example the description of \emph{authplugin} command below):

\begin{itemize}
\item \textit{inputcheck} - checks if all input files specified in job description
are downloadable.\\
\hspace*{0.5cm}inputcheck {[}-h] {[}-d debug\_level] RSL\_fle {[}proxy\_file]\\
\hspace*{0.5cm}- RSL\_file -- file with job description,\\
\hspace*{0.5cm}- proxy\_file -- credentials proxy.
\item \textit{lcas} - executes LCAS plugins on credentials and returns 0
if authorization passed.\\
\hspace*{0.5cm} lcas credentials description {[}library {[}db {[}directory]]]\\
\hspace*{0.5cm}- credentials -- path to file with credentials to
authorize,\\
\hspace*{0.5cm}- description -- path to file with job description,\\
\hspace*{0.5cm}- library -- path to LCAS library (full or relative
to LCAS directory),\\
\hspace*{0.5cm}- db -- path to LCAS DB file (full or relative to
LCAS directory),\\
\hspace*{0.5cm}- directory -- LCAS directory.
\end{itemize}

@@

The GM is installed into a single installation point referred as \$NORDUGRID\_LOCATION
and the following sub-directories are used:

\begin{LyXParagraphLeftIndent}{1cm}
\$NORDUGRID\_LOCATION/bin -- tools

\$NORDUGRID\_LOCATION/libexec -- program modules used by GM

\$NORDUGRID\_LOCATION/etc -- configuration files, deprecated, central
configuration file is used by default

>>>>

\begin{center}
\textsc{\large The ARC Computational Job Management Component - A-REX}
\par\end{center}{\large \par}

@@

A-REX is one of ARC middleware componets implementing functions of
so called \emph{Computing Element} (CE).

The aim of the A-REX is to take care of job pre- and post-processing.
It provides functionality to stage-in files containing input data
and program modules from wide range of sources and transfer or store
output results.

The A-REX implements Web Service (WS) interface which provides a way
to submit and control computational tasks (jobs) to be executed by
the A-REX and undelying Local Resource Management System.

@@

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files in the SD as well. The A-REX does not guarantee the
availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environement).
The SD is also the only place which is controlled by the A-REX. It
is accessible by the user from outside through HTTP(S) protocol. Any
file created outside the SD is not controlled by the A-REX. Any exchange
of data between client and A-REX (including also program modules)
is done through HTTP(S) protocol. A URL for accessing input/output
files is  obtained through the WS Local Information Description Interface
(LIDI) of A-REX.

Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
\cite{jsdl} XML document - which identifies the job in the A-REX
and the Information Interface.

Jobs are initiated and controlled through WS interface. Complete job
descriptions (JD)are passed to the A-REX through WS in JSDL\cite{jsdl}
coded description. Input data files and job executables are transferred
separately through the same interface, as described in the next section.

@@

One of the most important tasks of the A-REX is to take care of processing
input and output data (files) of the job. Input files are gathered
in the SD or associated cache area. There are two ways to put a file
into the SD:

\begin{itemize}
\item Download is initiated by the A-REX -- This is the case for files defined
in the JD (with name and source). The A-REX alone is responsible to
make sure that all required files will be available in the SD.\\
The supported protocols for sources at the moment are (in case of
full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
some less standard sources are supported. Those are described below.
The A-REX fully relies on HED framework \cite{hed} for data transfering
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-REX. Examples of such
files are the main executable of the job, the job's input files, etc.
These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and
any other remote data transport during execution of a job is not guaranteed
(at least not by A-REX).

@@

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service} (RLS) -- The A-REX takes care
of these files. They have to be specified in the JD. If the job fails
during any stage of processing, no attept is made to transfer those
files to their final destination, unless the option \emph{preserve=yes}
is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}
All files not specified in the JD are deleted after job execution
finished. If job execution fails for any reason (if exit code of main
executable is not 0) all files from furst group are transfered to
second one.

@@

From the point of view of the A-REX a job passes through various states.
Picture \ref{job states diagram} presents a diagram of the possible
states of a job.%
\begin{figure}[tbh]
\begin{centering}
\includegraphics{pic1}
\par\end{centering}

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the dedicated Local
Information Description Interface of A-REX using the UI or any other
suitable tool or through query method of WS interface.

@@

Below is the description of all actions taken by the A-REX at every
state:

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to a
CE but is not processed yet. The A-REX will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The A-REX is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If all files
are successfully gathered the job moves to the next state. If \textbf{any}
file can't be downloaded or it takes the UI too long to upload a file,
the job moves to \textbf{Finishing} state. It is possible to put a
limit on the number of simultaneous \textbf{Preparing} jobs. If this
limit is exceeded, jobs ready to enter the \textbf{Preparing} state
 will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
mark. Exceptions are jobs which have no files to be downloaded. These
are processed out of limits.
\item \textbf{Submiting} - The job is being passed for execution to the
\emph{Local Resource Management System} (LRMS). The correspoding backends
for many LRMSs are provided with the default installation. If the
local job submission is successful the job moves to the \textbf{Executing}
state. Otherwise it moves to \textbf{Finishing}. It is possible to
limit the agregate number of jobs in \textbf{Submiting} and \textbf{Executing}
states.
\item \textbf{Executing} - The job is queued or being executed in the LRMS.
The A-REX takes no actions except waiting until the job finishes.
\item \textbf{Killing} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at an Indexing Service. The user can download data files
from the SD by using the UI or other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the A-REX.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the Informatrion Interface of the A-REX. If a
job was moved to \textbf{Finished} because of failure, it may be restarted
on request of a client. When restarted, a job is moved to the state
previous to the one in which it failed and is assigned mark PENDING.
This is needed in order to not break the configuration limits. Exception
is a job failed in \textbf{Executing} state and lacking input files
specified in JD. Such a job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by
the user}. No files will be moved to their destination SE.

@@

In full installation the A-REX and it's components support the following
data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
http, https, lfc} and \emph{rls.} For more information please see
{}``The Hosting Environment of the Advanced Resource Connector middleware''
document \cite{hed}.

@@

\subsection{Internal Files of the A-REX}

For each local UNIX user listed in the A-REX configuration - including
a generic one which covers all local user identities - a \textit{control
directory} exists. In this directory the A-REX stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. In most common configuration case A-REX
serves all users defined by Operating System and stores their control
file in same directory. To make it easier to recover in case of failure,
the A-REX stores most information in files rather than in memory.
All files belonging to the same job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.description} - contains the description of the job
(JD).
\item \textit{job.ID.local} - information about the job used by the A-REX.
It consists of lines of format \textit{{}``name = value''}. Not
all of them are always available. The following names are defined:

@@

\item \textit{job.ID.proxy} -delegated X509 credentials.
\item \textit{job.ID.proxy.tmp} - temporary X509 credentials with different
UNIX ownership used by processes run with effective \emph{user id}
different form job owner's \emph{id}.
\end{itemize}
There are other files with names like job.ID.{*} which are created
and used by different parts of the A-REX. Their presence in the \textit{control
directory} can not be guaranteed and can change depending on changes
in the A-REX code.

@@

The job submission and control interface is based on document produced
by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
The exchange of SOAP messages is performed over the HTTP(S) protocol.
The BES interface is represented by two port-types - BES-Management
and BES-Factory. The former is made to control the A-REX service itself
and thus defines operations to start and stop the functionality of
the BES service. The A-REX does not implement remote control of service
functionality. Hence the BES-Management port-type is not functional.
The BES-Factory port-type provides operations to submit new jobs (to
create an activity in terms of BES) and to monitor its state. It also
has an ability to provide information about the service. A-REX fully
implements the functionality of this port-type.

For job descriptions A-REX accepts the Job Submission Description
Language (JSDL)\cite{jsdl} documents as defined by the OGF Job Submission
Description Language Working Group. Supported elements and extensions
are described below.

@@

A-REX introduces a new operation in addition to those provided by
BES. It does that by defining its own port-type with the single operation
\emph{ChangeActivityStatus} (see Appendix D). This operation provides
a way to request simple transfers between states of jobs and corresponding
actions.

@@

Since BES allows implementators to extend their initial activity states
with additional sub-states, A-REX defines a set of sub-states of activity
processing in addition to those defined by the BES, as listed in Table
\ref{tab:Job-states-definitions}. Their meaning is described in section
\ref{Section: Job Flow}.

@@

The A-REX also supportd the Delegation Interface (see Appendix E).
This is a common purpose interface to be used by ARC services which
accepts delegated credentials from clients. The Delegation Interface
implements two operations: initiatialization of credentials delegation
(DelegateCredentialsInit) and update/renewal of credentials (UpdateCredentials).

@@

\begin{itemize}
\item \emph{UnsupportedFault}: Indicates that service does not support this
operation despite supporting the port-type.
\item \emph{ProcessingFault}: Internal problems during generation of the
token.
\end{itemize}
\end{itemize}
\end{itemize}
Additionally, A-REX Web Service Interface allows delegation to be
performed as part of the \emph{CreateActivity} operation of the BES-Factory
port-type. For this it accepts the element \emph{DelegatedCredentials}
inside the \emph{CreateActivity} element. The \emph{Id} element of
\emph{DelegatedCredentials} must contain an identifier obtained in
response to the previous \emph{DelegateCredentialsInit} operation.
For more information about delegations and delegation interface refer
to \cite{wsrf-rp}.

@@

The A-REX implements the Local Information Description Interface (LIDI)
interface common for all ARC services. This interface is based on
OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
Information about resources and maintained activities/jobs are represented
in a \emph{WS-Resource Properties} informational XML document. The
document type is defined in the A-REX WSDL as a \emph{ResourceInformationDocumentType}.
It contains the following elements/resources:

@@

A-REX supports the following elements from the JSDL version 1.0 specification\cite{jsdl}
including POSIX Applications extension and JSDL HPC Profile Application
Extension {[}]:

\begin{lyxlist}{00.00.0000}
\item [{\emph{JobName}}] \emph{-} name of the job as assigned by the user.
\item [{\emph{Executable}\ (POSIX,HPC)}] - name of the executable file.
\item [{\emph{Argument}\ (POSIX,HPC)}] - arguments the executable will
be launched with.
\item [{\emph{DataStaging}}]~

\begin{lyxlist}{00.00.0000}
\item [{Filename}] - name of the data file on the executing node.
\item [{\emph{Source}}] - source where the file will be taken from before
execution.
\item [{\emph{Target}}] \emph{-} destination the file will be delivered
to after execution.
\end{lyxlist}
\item [{\emph{Input}\ (POSIX,HPC)}] - file to be used as standard input
for the executable.
\item [{\emph{Output}\ (POSIX,HPC)}] - file to be used as standard output
for the executable.
\item [{\emph{Error}\ (POSIX,HPC)}] - file to be used as standard error
for the executable.
\item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
for execution.
\item [{\emph{TotalPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{IndividualPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{CPUTimeLimit}\ (POSIX)}] - maximal amount of CPU time needed
for execution.
\item [{\emph{TotalCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{IndividualCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{WallTimeLimit}\ (POSIX)}] - amount of clock time needed
for execution.
\item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
\item [{\emph{IndividualCPUCount}}] - same as TotalCPUCount.
\item [{\emph{Environment}\ (POSIX)}] - environement variable.
\end{lyxlist}

@@

The A-REX can cache input files. Caching is enabled if corresponding
command is present in configuration file. The A-REX does not cache
files marked as executable in job. Caching can also be explicitly
turned off by user for each file by using \textit{cache=no} option
in URL (for URL options read {}``Protocols, Uniform Resource Locators
(URL) and extensions supported in ARC'' \cite{urls}). The disc space
occupied by cache is controlled by removing unused files. For more
information look in section \ref{SubSection: ConfigFile}.

\emph{NOTE: The caching algorithm used by A-REX and described above
is currently being replaced by a better one which is no robust and
provides better scalability.}

@@

Before downloading a file the A-REX tries to determine it's size and
then tries to preallocate space in the cache directory by writing
an empty file of the same size. If this fails (possibly because the
file system has no more space), it tries to the remove oldest cached
files which are not being used by any job. This means that \textbf{the
real hard limit of cache size is the space available in the file-system}.
In case cache gets full and it is impossible to free enough space,
the download fails and then is retried without using cache.

Before giving access to an already cached file, the A-REX contacts
the original file source to check if the user has suficient access
rights to the file. Not all protocols make this check possible. Also,
file creation or validity times are checked to make sure the cached
file is fresh. If it is impossible to obtain creation and invalidation
times, the file is invalidated 24 hours after it was downloaded.

The A-REX checks the cache periodically. If the space used by cache
exceeds the high water-mark given in the configuration file (\emph{softsize}),
it tries to remove the oldest unused files until cache size drops
below the low water-mark. This sets the soft limit of cache size.

There are two kinds of caches supported. Files in \textit{private}
cache are owned by the Unix user to which a Grid user is mapped. Those
files are readable only by that particular Unix user. Another kind
of cache is \textit{shared}. Files in the shared cache are owned by
the Unix user who started A-REX and are readable by everyone.

@@

The A-REX consists of several separate modules. These are:

\begin{itemize}
\item \textit{libarex.so} - The main module providing main functionality
and web interface. It is implemented as HTTP and SOAP service inside
HED. It is responsible for processing jobs, moving them through states
and running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at an Indexding Service
(like RLS) as needed. It processes and updates the \textit{job.ID.output}
file.
\item \emph{cache-register} - Utility to register cached data into an Indexing
Service. It reads and modifies cache informational files \emph{old}
and \emph{new} (as described in section \ref{sec:cache}). Configuration
is read directly from the GM's configuration file (see section \ref{SubSection: ConfigFile}).
It is run by the GM every 5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
the frontend. It puts collected information into the \emph{job.ID.diag}
file.
\item \emph{gm-kick} - Sends a signal to the A-REX though a FIFO file to
wake it up. It's used to increase responsiveness of A-REX.
\item \emph{cluster+qju.pl} - Collects and generates information about computing
resource as XML document in NoruGrid and Glue 2 format.
\end{itemize}
The following modules are always run under the Unix account to which
a Grid user is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Curently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
pseudo-LRMS is supported for testing purposes. This module is responsible
for job submission to the LRMS.
\item \textit{cancel-{*}-job} - This script is for canceling jobs which
have been already submitted to the LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the A-REX about completion of jobs. It's implementation for PBS uses
server logs to extract information about jobs. If logs are not available
it uses the less reliable \emph{qstat} command for that. Other backends
use different techniques.
\end{itemize}
In addition, there is also an administration utility:

\begin{itemize}
\item \textit{gm-jobs} - prints a list of jobs available on the cluster
and the number of jobs in each state.\\
\hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name]\\
\hspace*{0.5cm}-l -- print more information about each job,\\
\hspace*{0.5cm}-u -- pretend utility is run by user with id \emph{uid},\\
\hspace*{0.5cm}-U -- pretend utility is run by user with name \emph{name}.
\end{itemize}

@@

The A-REX is installed into a single installation point referred as
\$ARC\_LOCATION and the following sub-directories are used:

\begin{LyXParagraphLeftIndent}{1cm}
\$ARC\_LOCATION/bin -- tools

\$ARC\_LOCATION/libexec -- program modules used by A-REX

/etc -- central configuration file - location used by default

\$ARC\_LOCATION/lib/arc -- service module
